---
title: "02-Intermediate-R-for-Finance"
author: "Boni"
date: "9/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working with Date Time
```{r}
# Today
x <- Sys.Date()
y <- Sys.time()
x
y
class(x) # Calendar date class
class(y) # POSIX class, with POSIXct and POSIXlt

# Convert string to Date
date <- "2020-09-01" # Use the standard ISO 8601 Standard: year-month-day or year/month/day
date_as_date <- as.Date(date)
class(date_as_date)
as.numeric(date_as_date) # Find number of days since January 1, 1970

# Explicit conversion
as.Date("May 2020 09", format = "%b %Y %d")

# char_dates
char_dates <- c("1jan17", "2jan17", "3jan17", "4jan17", "5jan17")

# Create dates using as.Date() and the correct format 
dates <- as.Date(char_dates, format="%d%b%y")

# Use format() to go from "2017-01-04" -> "Jan 04, 17"
format(dates, format = "%b %d, %y")

# Use format() to go from "2017-01-04" -> "01,04,2017"
format(dates, format = "%m,%d,%Y")

# Dates
dates <- as.Date(c("2017-01-01", "2017-01-02", "2017-01-03"))

# Create the origin
origin <- as.Date("1970-01-01")

# Use as.numeric() on dates
as.numeric(dates)

# Find the difference between dates and origin
dates - origin
```
## Extract information from Date Time
```{r}
# dates
dates <- as.Date(c("2017-01-02", "2017-05-03", "2017-08-04", "2017-10-17"))

# Extract the months
months(dates)

# Extract the quarters
quarters(dates)

# dates2
dates2 <- as.Date(c("2017-01-02", "2017-01-03", "2017-01-04", "2017-01-05"))

# Assign the weekdays() of dates2 as the names()
names(dates2) <- weekdays(dates2)

# Print dates2
dates2
```

## Relational operators
```{r}
today <- 23.45
yesterday <- 34.23
today > yesterday

one <- 1
one == TRUE
```

## Working with loops
```{r}
# Using repeat
i <- 0

repeat {
  i <- i + 1
  print(i)
  if (i == 9) {
    break
  }
}

# while loop
stock_price <- 53.89

while (stock_price < 60) {
  stock_price <- stock_price * runif(1, 0.8, 1.2)
  print(stock_price)
}

# For loops
x <- 1:5
for (n in x) {
  print(n)
}
```

## Working with function
```{r}
x <- runif(9, 0.3, 1.4)
x
diff(x) # X_i - X_(i-t)

# Arithmetic return
prices <- c(23.4, 23.8, 22.3)

arith_return <- function(x) {
  diff(x) / x[-length(x)]
}

arith_return(prices) # (X_i - X_(i-t)) / X_(i-t)
```

## Functions and packages related to financial data
```{r}
library(tidyquant)
library(quantmod)

quantmod::getSymbols(c("IBM", "MSFT"))
plot(IBM$IBM.Adjusted)
plot(MSFT$MSFT.Adjusted)
```

```{r}
# Pull Apple stock data
apple <- tq_get("AAPL", get = "stock.prices", 
                from = "2007-01-03", to = "2017-06-05")

# Take a look at what it returned
head(apple)

# Plot the stock price over time
plot(apple$date, apple$adjusted, type = "l")

# Calculate daily stock returns for the adjusted price
apple <- tq_mutate(data = apple,
                   ohlc_fun = Ad,
                   mutate_fun = dailyReturn)

# Sort the returns from least to greatest
sorted_returns <- sort(apple$daily.returns)

# Plot them
plot(sorted_returns)
```

## Use apply function
There are many apply family:

- apply (Apply function over array margins)
- **lapply** (Apply a function over a list or vector)
- eapply (Apply a function over values in environment)
- mapply (Apply a function to multiple lists or vector arguments)
- rapply (Recursively apply a function to a list)
- tapply (Apply a function over ragged array)
- **sapply** (Simplify the result from lapply)
- **vapply** (Strictly simplify the result from lapply)
```{r}
# lapply in list
list1 <- list("Boni", 31, "University of Chicago", TRUE)
names(list1) <- c("name", "age", "university", "handsome")
lapply(list1, class)
# lapply in data frame
a <- c(1,2)
b <- c(3,4)
df <- data.frame(a, b)
lapply(df, sum) # lapply will create summary of each columns
```
Many times lapply works better than sapply because sapply simplify many things that might be not expected.
```{r}
# sapply in list
list1 <- list("Boni", 31, "University of Chicago", TRUE)
names(list1) <- c("name", "age", "university", "handsome")
sapply(list1, class)

# sapply in data frame
a <- c(1,2)
b <- c(3,4)
df <- data.frame(a, b)
sapply(df, sum) # sapply will create summary of each columns
```

```{r}
# Market crash with as.POSIXct()
market_crash2 <- list(dow_jones_drop = 777.68, 
                      date = as.POSIXct("2008-09-28"))

# Find the classes with sapply()
sapply(market_crash2, class)

# Find the classes with vapply()
# vapply(market_crash2, class, FUN.VALUE = character(1))

df <- data.frame(ibm = runif(10, 3, 9), apple = rnorm(10, 3, 1))
vapply(df, FUN = function(x) {c(max = max(x), min = min(x))}, FUN.VALUE = numeric(2))
```

