---
title: "Manipulating Time Series Data with xts and zoo in R"
author: "Boni"
date: "9/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Time Series Data with xts and zoo
xts stands for Extensible Time Series

```{r}
library(xts)
library(zoo)
# XTS = MATRIX + INDEX
x <- matrix(1:6, ncol = 2, nrow = 3)
x

idx <- as.Date(c("2020-05-09", "1989-05-09", "2019-09-17"))
idx

X <- xts(x = x, order.by = idx) # R will assign the DF with the time data, and reorder the DF based on the date assigned to it.
X
```

```{r}
# Create the object data using 5 random numbers
data <- rnorm(5)

# Create dates as a Date class object starting from 2016-01-01
dates <- seq(as.Date("2016-01-01"), length = 5, by = "days")

# Use xts() to create smith
smith <- xts(x = data, order.by = dates)

# Create bday (1899-05-08) using a POSIXct date class object
bday <- as.POSIXct("1899-05-08")

# Create hayek and add a new attribute called born
hayek <- xts(x = data, order.by = dates, born = bday)

attributes(smith)
attributes(hayek)
```
```{r}
# Extract the core data of hayek
hayek_core <- coredata(hayek)

# View the class of hayek_core
class(hayek_core)

# Extract the index of hayek
hayek_index <- index(hayek)

# View the class of hayek_index
class(hayek_index)
```

```{r}
# Create dates
dates <- as.Date("2016-01-01") + 0:4

# Create ts_a
ts_a <- xts(x = 1:5, order.by = dates)

# Create ts_b
ts_b <- xts(x = 1:5, order.by = as.POSIXct(dates))

# Extract the rows of ts_a using the index of ts_b
ts_a[index(ts_b)]

# Extract the rows of ts_b using the index of ts_a
ts_b[index(ts_a)]
```

```{r}
data("sunspots")
class(sunspots)
head(sunspots)

sunspots_xts <- as.xts(sunspots)
head(sunspots_xts)
```

```{r}
# Convert austres to an xts object called au
au <- as.xts(austres)

# Then convert your xts object (au) into a matrix am
am <- as.matrix(au)

# Inspect the head of am
head(am)

# Convert the original austres into a matrix am2
am2 <- as.matrix(austres)

# Inspect the head of am2
head(am2)

```
```{r}
first(sunspots_xts, "5 months")
last(sunspots_xts, "2 years")
```
**notes**

- Use rbind to add rows from multiple xts
- Use join to merge multiple xts with different column
- Use na.locf() to fill missing value with the previous date value. Add parameter fromLast to fill the missing value with the next date value.
- Use na.approx() to fill missing value with interpolate value using linear approximation.
- Use split() to split the data by week, month, year, etc.
- Use to.period() to get OHLC information by month, year, week, etc.
- Use index() to get the time index.
- Use indexClass() to get the index class.
- Use indexFormat() to get or set the index time format.
- Use tzone() to get the time zone of the index.