---
title: "Time Series Analysis in R"
author: "Boni"
date: "9/4/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Nile dataset
```{r}
data("Nile")
```
# Explore Nile
```{r}
# Print the Nile dataset
print(Nile)

# List the number of observations in the Nile dataset
length(Nile)

# Display the first 10 elements of the Nile dataset
head(Nile, 10)

# Display the last 12 elements of the Nile dataset
tail(Nile, 12)
```

# Plot time series data
```{r}
# Plot the Nile data
plot(Nile)

# Plot the Nile data with xlab and ylab arguments
plot(Nile, xlab = "Year", ylab = "River Volume (1e9 m^{3})")

# Plot the Nile data with xlab, ylab, main, and type arguments
plot(Nile, xlab = "Year", ylab = "River Volume (1e9 m^{3})", main = "Annual River Nile Volume at Aswan, 1871-1970", type = "b")
```

# Identifying time series frequency

The start() and end() functions return the time index of the first and last observations, respectively. The time() function calculates a vector of time indices, with one element for each time index on which the series was observed.

The deltat() function returns the fixed time interval between observations and the frequency() function returns the number of observations per unit time. Finally, the cycle() function returns the position in the cycle of each observation.

```{r}
# Plot AirPassengers
plot(AirPassengers)

# View the start and end dates of AirPassengers
start(AirPassengers)
end(AirPassengers)

# Use time(), deltat(), frequency(), and cycle() with AirPassengers 
time(AirPassengers)
deltat(AirPassengers)
frequency(AirPassengers)
cycle(AirPassengers)

```

# Creating ts object
```{r}
data_vector <- rnorm(50, 3, 2.6)
# Use print() and plot() to view data_vector
print(data_vector)
plot(data_vector)

# Convert data_vector to a ts object with start = 2004 and frequency = 4
time_series <- ts(data_vector, start = 2004, frequency = 4)

# Use print() and plot() to view time_series
print(time_series)
plot(time_series)

```

# Check wheter a variable is time series
```{r}
# Check whether data_vector and time_series are ts objects
is.ts(data_vector)
is.ts(time_series)

# Check whether Nile is a ts object
is.ts(Nile)

# Check whether AirPassengers is a ts object
is.ts(AirPassengers)
```

# Plotting time series object
```{r}
eu_stocks <- EuStockMarkets
# Check whether eu_stocks is a ts object
is.ts(eu_stocks)

# View the start, end, and frequency of eu_stocks
start(eu_stocks)
end(eu_stocks)
frequency(eu_stocks)

# Generate a simple plot of eu_stocks
plot(eu_stocks)

# Use ts.plot with eu_stocks
ts.plot(eu_stocks, col = 1:4, xlab = "Year", ylab = "Index Value", main = "Major European Stock Indices, 1991-1998")

# Add a legend to your ts.plot
legend("topleft", colnames(eu_stocks), lty = 1, col = 1:4, bty = "n")

```

# Trend spotting

## Removing trends in variability via the logarithmic transformation

The logarithmic function log() is a data transformation that can be applied to positively valued time series data. It slightly shrinks observations that are greater than one towards zero, while greatly shrinking very large observations. This property can stabilize variability when a series exhibits increasing variability over time. It may also be used to linearize a rapid growth pattern over time.

```{r}
colnames(eu_stocks)
smi_rapid_growth <- eu_stocks[,1]
smi_linear_growth <- log(smi_rapid_growth)
ts.plot(smi_linear_growth)
ts.plot(smi_rapid_growth)
```

## Removing trends in level by differencing
The first difference transformation of a time series z[t] consists of the differences (changes) between successive observations over time, that is z[t] - z[t-1].

Differencing a time series can remove a time trend. The function diff() will calculate the first difference or change series. A difference series lets us examine the increments or changes in a given time series. It always has one fewer observations than the original series.

By removing the long-term time trend, we can view the amount of change from one observation to the next.
```{r}
dsmi <- diff(smi_rapid_growth)
length(smi_rapid_growth)
length(dsmi)
ts.plot(dsmi)
```

## Removing seasonal trends with seasonal differencing

For time series exhibiting seasonal trends, seasonal differencing can be applied to remove these periodic patterns. For example, monthly data may exhibit a strong twelve month pattern. In such situations, changes in behavior from year to year may be of more interest than changes from month to month, which may largely follow the overall seasonal pattern.

The function diff(..., lag = s) will calculate the lag s difference or length s seasonal change series. For monthly or quarterly data, an appropriate value of s would be 12 or 4, respectively. The diff() function has lag = 1 as its default for first differencing. Similar to before, a seasonally differenced series will have s fewer observations than the original series.

```{r}
# create x
x <- rpois(100, 3)
# Generate a diff of x with lag = 4. Save this to dx
dx <- diff(x, lag = 2)
  
# Plot dx
ts.plot(dx)

# View the length of x and dx, respectively 
length(x)
length(dx)

```

## Simulate the white noise model
The white noise (WN) model is a basic time series model. We will focus on the simplest form of WN, independent and identically distributed data.

The arima.sim() function can be used to simulate data from a variety of time series models. ARIMA is an abbreviation for the autoregressive integrated moving average class of models.

An ARIMA(p, d, q) model has three parts, the autoregressive order p, the order of integration (or differencing) d, and the moving average order q. 
```{r}
# Simulate a WN model with list(order = c(0, 0, 0))
white_noise <- arima.sim(model = list(order = c(0,0,0)), n = 100)

# Plot your white_noise data
ts.plot(white_noise)

# Simulate from the WN model with: mean = 100, sd = 10
white_noise_2 <- arima.sim(model = list(order = c(0,0,0)), n = 100, mean = 100, sd = 10)

# Plot your white_noise_2 data
ts.plot(white_noise_2)

```

## Estimate the white noise model

For a given time series y we can fit the white noise (WN) model using the arima(..., order = c(0, 0, 0)) function. Recall that the WN model is an ARIMA(0,0,0) model. Applying the arima() function returns information or output about the estimated model. For the WN model this includes the estimated mean, labeled intercept, and the estimated variance, labeled sigma ^ 2.

```{r}
y <- white_noise
# Fit the WN model to y using the arima command
arima(y, order = c(0, 0, 0))

# Calculate the sample mean and sample variance of y
mean(y)
var(y)

```

## Simulate the random walk model
The random walk (RW) model is also a basic time series model. It is the cumulative sum (or integration) of a mean zero white noise (WN) series, such that the first difference series of a RW is a WN series. Note for reference that the RW model is an ARIMA(0, 1, 0) model, in which the middle entry of 1 indicates that the model's order of integration is 1.

The arima.sim() function can be used to simulate data from the RW by including the model = list(order = c(0, 1, 0)) argument. We also need to specify a series length n. Finally, you can specify a sd for the series (increments), where the default value is 1.

```{r}
# Generate a RW model using arima.sim
random_walk <- arima.sim(model = list(order = c(0, 1, 0)), n = 100)

# Plot random_walk
ts.plot(random_walk)

# Calculate the first difference series
random_walk_diff <- diff(random_walk)

# Plot random_walk_diff
ts.plot(random_walk_diff)
  
```

## Simulate the random walk model with a drift
A random walk (RW) need not wander about zero, it can have an upward or downward trajectory, i.e., a drift or time trend. This is done by including an intercept in the RW model, which corresponds to the slope of the RW time trend.

For an alternative formulation, you can take the cumulative sum of a constant mean white noise (WN) series, such that the mean corresponds to the slope of the RW time trend.

To simulate data from the RW model with a drift you again use the arima.sim() function with the model = list(order = c(0, 1, 0)) argument. This time, you should add the additional argument mean = ... to specify the drift variable, or the intercept.

```{r}
# Generate a RW model with a drift uing arima.sim
rw_drift <- arima.sim(model = list(order = c(0, 1, 0)), n = 100, mean = 1)

# Plot rw_drift
ts.plot(rw_drift)

# Calculate the first difference series
rw_drift_diff <- diff(rw_drift)

# Plot rw_drift_diff
ts.plot(rw_drift_diff)

```

## Estimate the random walk model
For a given time series y we can fit the random walk model with a drift by first differencing the data, then fitting the white noise (WN) model to the differenced data using the arima() command with the order = c(0, 0, 0)) argument.

The arima() command displays information or output about the fitted model. Under the Coefficients: heading is the estimated drift variable, named the intercept. Its approximate standard error (or s.e.) is provided directly below it. The variance of the WN part of the model is also estimated under the label sigma ^ 2.

```{r}
# Difference your random_walk data
rw_diff <- diff(random_walk)

# Plot rw_diff
ts.plot(rw_diff)

# Now fit the WN model to the differenced data
model_wn <-arima(rw_diff, order = c(0, 0, 0))

# Store the value of the estimated time trend (intercept)
int_wn <- model_wn$coef

# Plot the original random_walk data
ts.plot(random_walk)

# Use abline(0, ...) to add time trend to the figure
abline(0, int_wn)

```

## Are the white noise model or the random walk model stationary?

The white noise (WN) and random walk (RW) models are very closely related. However, only the RW is always non-stationary, both with and without a drift term. This is a simulation exercise to highlight the differences.

Recall that if we start with a mean zero WN process and compute its running or cumulative sum, the result is a RW process. The cumsum() function will make this transformation for you. Similarly, if we create a WN process, but change its mean from zero, and then compute its cumulative sum, the result is a RW process with a drift.

```{r}
# Use arima.sim() to generate WN data
white_noise <- arima.sim(model = list(order = c(0, 0, 0)), n = 100)

# Use cumsum() to convert your WN data to RW
random_walk <- cumsum(white_noise)
  
# Use arima.sim() to generate WN drift data
wn_drift <- arima.sim(model = list(order = c(0, 0, 0)), n = 100, mean = 0.4)
  
# Use cumsum() to convert your WN drift data to RW
rw_drift <- cumsum(wn_drift)

# Plot all four data objects
plot.ts(cbind(white_noise, random_walk, wn_drift, rw_drift))

```

## Asset prices vs. asset returns

The goal of investing is to make a profit. The revenue or loss from investing depends on the amount invested and changes in prices, and high revenue relative to the size of an investment is of central interest. This is what financial asset returns measure, changes in price as a fraction of the initial price over a given time horizon, for example, one business day.

Let's again consider the eu_stocks dataset. This dataset reports index values, which we can regard as prices. The indices are not investable assets themselves, but there are many investable financial assets that closely track major market indices, including mutual funds and exchange traded funds.

Log returns, also called continuously compounded returns, are also commonly used in financial time series analysis. They are the log of gross returns, or equivalently, the changes (or first differences) in the logarithm of prices.

The change in appearance between daily prices and daily returns is typically substantial, while the difference between daily returns and log returns is usually small. As you'll see later, one advantage of using log returns is that calculating multi-period returns from individual periods is greatly simplified - you just add them together!

```{r}
# Plot eu_stocks
plot(eu_stocks)

# Use this code to convert prices to returns
returns <- eu_stocks[-1,] / eu_stocks[-1860,] - 1

# Convert returns to ts
returns <- ts(returns, start = c(1991, 130), frequency = 260)

# Plot returns
plot(returns)

# Use this code to convert prices to log returns
logreturns <- diff(log(eu_stocks))

# Plot logreturns
plot(logreturns)

```

## Characteristics of financial time series

Daily financial asset returns typically share many characteristics. Returns over one day are typically small, and their average is close to zero. At the same time, their variances and standard deviations can be relatively large. Over the course of a few years, several very large returns (in magnitude) are typically observed. These relative outliers happen on only a handful of days, but they account for the most substantial movements in asset prices. Because of these extreme returns, the distribution of daily asset returns is not normal, but heavy-tailed, and sometimes skewed. In general, individual stock returns typically have even greater variability and more extreme observations than index returns.

```{r}
eu_percentreturns <- diff(eu_stocks)[,] / eu_stocks[1:length(diff(eu_stocks))] * 100
head(eu_percentreturns)
# Generate means from eu_percentreturns
colMeans(eu_percentreturns)

# Use apply to calculate sample variance from eu_percentreturns
apply(eu_percentreturns, MARGIN = 2, FUN = var)

# Use apply to calculate standard deviation from eu_percentreturns
apply(eu_percentreturns, MARGIN = 2, FUN = sd)

# Display a histogram of percent returns for each index
par(mfrow = c(2,2))
apply(eu_percentreturns, MARGIN = 2, FUN = hist, main = "", xlab = "Percentage Return")

# Display normal quantile plots of percent returns for each index
par(mfrow = c(2,2))
# apply(eu_percentreturns, MARGIN = 2, FUN = qqnorm, main = "")
qqline(eu_percentreturns)

```

## Plotting pairs of data

Time series data is often presented in a time series plot. For example, the index values from the eu_stocks dataset are shown in the adjoining figure. Recall, eu_stocks contains daily closing prices from 1991-1998 for the major stock indices in Germany (DAX), Switzerland (SMI), France (CAC), and the UK (FTSE).

It is also useful to examine the bivariate relationship between pairs of time series. In this exercise we will consider the contemporaneous relationship, that is matching observations that occur at the same time, between pairs of index values as well as their log returns. The plot(a, b) function will produce a scatterplot when two time series names a and b are given as input.

To simultaneously make scatterplots for all pairs of several assets the pairs() function can be applied to produce a scatterplot matrix. When shared time trends are present in prices or index values it is common to instead compare their returns or log returns.

```{r}
# Make a scatterplot matrix of eu_stocks
pairs(eu_stocks)

# Convert eu_stocks to log returns
logreturns <- diff(log(eu_stocks))

# Plot logreturns
plot(logreturns)

# Make a scatterplot matrix of logreturns
pairs(logreturns)

```

## 